\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-06 paper.qmd
Updated figure 17 title

By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout. As shown in @mnist-model-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$. The model shows some quirks. The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@@mnist-model-proj (b)). Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@@mnist-model-proj (c)).

@@

<!--add langevitour screenshots and youtube animation link-->
```{r}
#| label: fig-mnist-tri-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.062$, $b = 285 \\  (19, \ 15)$, $m = 109$, and benchmark value to remove large edges is $0.108$). (b) The non-linear structure observed in the \\gD{} layout of PaCMAP (@fig-pacmap-author) is also visible when visualizing the model overlaid on the data space. This indicates that PaCMAP accurately captures the non-linear structure of the $10\\text{-}D$ data. (c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-pacmap-author). (d) some long edges exist in the $10\\text{-}D$ space that are not recognized as long edges in the \\gD{} representation. The full video is available at <https://youtu.be/zcg_GXBmqjA>."

>>>>

By visualizing the model generated for tSNE in $10\text{-}D$ helps to assess the \gD{} layout. As shown in @fig-mnist-tri-proj (a), the model provides the evidence for the non-linear structure of the digit 1 data in $10\text{-}D$. The model shows some quirks. The model's twisted pattern provides evidence for the $10\text{-}D$ data structure, which is not observed by \gD{} layout (@fig-mnist-tri-proj (b)). Furthermore, the presence of long edges indicates the existence of the small cluster located closely together (@fig-mnist-tri-proj (c)).

@@

<!--add langevitour screenshots and youtube animation link-->
```{r}
#| label: fig-mnist-tri-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the tSNE layout (a), and three views (b, c, d) of the fit in projections from $10\\text{-}D$, for the digit 1 of MNIST data ($a_1 = 0.07$, $b = 456 \\  (19, \ 24)$, $m = 226$, and benchmark value to remove large edges is $0.114$). (b) The non-linear structure observed in the \\gD{} layout of tSNE (@fig-tsne-best) is also visible when visualizing the model overlaid on the data space. This indicates that tSNE accurately captures the non-linear structure of the $10\\text{-}D$ data. (c) the model shows a twisted pattern within the non-linear structure in $10\\text{-}D$ space, which is an additional pattern not visible in the \\gD{} representation (@fig-tsne-best). (d) some long edges exist in the $10\\text{-}D$ space that are not recognized as long edges in the \\gD{} representation. The full video is available at <https://youtu.be/zcg_GXBmqjA>."

### 2024-09-06 paper.tex
Updated figure 17 title

By visualizing the model generated for tSNE in \(10\text{-}D\) helps to
assess the \gD{} layout. As shown in \citet{mnist-model-proj} (a), the
model provides the evidence for the non-linear structure of the digit 1
data in \(10\text{-}D\). The model shows some quirks. The model's
twisted pattern provides evidence for the \(10\text{-}D\) data
structure, which is not observed by \gD{} layout
(@\citet{mnist-model-proj} (b)). Furthermore, the presence of long edges
indicates the existence of the small cluster located closely together
(@\citet{mnist-model-proj} (c)).

@@

\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the tSNE layout
(a), and three views (b, c, d) of the fit in projections from
\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.062\),
\(b = 285 \  (19,  15)\), \(m = 109\), and benchmark value to remove
large edges is \(0.108\)). (b) The non-linear structure observed in the
\gD{} layout of PaCMAP (\textbf{?@fig-pacmap-author}) is also visible
when visualizing the model overlaid on the data space. This indicates
that PaCMAP accurately captures the non-linear structure of the
\(10\text{-}D\) data. (c) the model shows a twisted pattern within the
non-linear structure in \(10\text{-}D\) space, which is an additional
pattern not visible in the \gD{} representation
(\textbf{?@fig-pacmap-author}). (d) some long edges exist in the
\(10\text{-}D\) space that are not recognized as long edges in the \gD{}
representation. The full video is available at
\url{https://youtu.be/zcg_GXBmqjA}.}

>>>>

By visualizing the model generated for tSNE in \(10\text{-}D\) helps to
assess the \gD{} layout. As shown in Figure~\ref{fig-mnist-tri-proj}
(a), the model provides the evidence for the non-linear structure of the
digit 1 data in \(10\text{-}D\). The model shows some quirks. The
model's twisted pattern provides evidence for the \(10\text{-}D\) data
structure, which is not observed by \gD{} layout
(Figure~\ref{fig-mnist-tri-proj} (b)). Furthermore, the presence of long
edges indicates the existence of the small cluster located closely
together (Figure~\ref{fig-mnist-tri-proj} (c)).

@@

\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the tSNE layout
(a), and three views (b, c, d) of the fit in projections from
\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.07\),
\(b = 456 \  (19,  24)\), \(m = 226\), and benchmark value to remove
large edges is \(0.114\)). (b) The non-linear structure observed in the
\gD{} layout of tSNE (Figure~\ref{fig-tsne-best}) is also visible when
visualizing the model overlaid on the data space. This indicates that
tSNE accurately captures the non-linear structure of the \(10\text{-}D\)
data. (c) the model shows a twisted pattern within the non-linear
structure in \(10\text{-}D\) space, which is an additional pattern not
visible in the \gD{} representation (Figure~\ref{fig-tsne-best}). (d)
some long edges exist in the \(10\text{-}D\) space that are not
recognized as long edges in the \gD{} representation. The full video is
available at \url{https://youtu.be/zcg_GXBmqjA}.}
